---
title: "RGraphics"
author: "Sanjeev Kumar"
date: "Oct 23, 2019"
output: 
  html_document: 
    toc: yes
    toc_float: yes
---

> Today we will explore Data Visualization in R using the ggplot2 package. We will also do some Geographical Visualization. Time permitting we will explore the dplyr package for data manipulation.

## ggplot2

Let's begin by loading the package
```{r}
library("ggplot2") #Assuming that you have already installed the package
```

There are two main functions in ggplot2: `qplot()` and `ggplot()`. We will focus exclusively on `ggplot()` as `qplot()` is just a traditional `plot()` like implementation of `ggplot()`.

> **Key Concept of ggplot**: Data Visualization is about mapping properties of data to visual properties in the graphics. `ggplot` allows you full freesom to map *any* data property to *any* visual property.

### Key Elements of `ggplot`
- **Data**: What you want to visualize. Consists of `variables` stored in columns of a data frame
- **Geoms**: Geometric Objects drawn to represent the data such as lines, points, bars etc.
- **Aesthetics**: Visual properties of Geoms such as line color, point shapes, x-y position
- **Mappings** of data values to Aesthetics
- **Scales** control the mappings
- **Guides** show how to map visual properties back to data - for example - Labels

### Test Drawings

We will use the `diamonds` dataset that is built into `ggplot2`. Lets take a quick look into the data:
```{r}
dim(diamonds)
head(diamonds)
```

Lets start by making a simple scatter plot of `carat` vs `price`
```{r}
baseplot <- ggplot(data=diamonds, aes(x=carat, y=price))
```

The basic `ggplot` function does not draw anything. It just creates the infrastructure based on your data. We can now add the `geom`
```{r}
baseplot + geom_point()
```

Arguments passed to `aes()` are the mappings. Further specifications can be done as arguments to the `geom()` function call.
```{r}
baseplotcolor <- ggplot(data=diamonds, aes(x=carat, y=price, colour=color))
baseplotcolor + geom_point(size=3, alpha=1)
```

Perhaps you want y axis on log scale
```{r}
baseplotcolor + geom_point(size=0.7, alpha=0.3) + scale_y_log10()
```

We can do the same chart for different values of `cut`
```{r}
baseplotcolor + geom_point(size=0.7, alpha=0.3) + facet_wrap(~ cut)
```

### Different Tyoes of Geoms

`ggplot` can create a lot of different kinds of plots.  

Function | Description
---------|------------
`geom_point()` | Points, i.e., scatterplot
`geom_bar()` | Bar chart
`geom_line()` | Line chart
`geom_boxplot()` | Boxplot
`geom_violin()` | Violin plot
`geom_density()` | Density plot with one variable
`geom_density2d()` | Density plot with two variables
`geom_histogram()` | Histogram

For example: **bar charts**
```{r}
ggplot(data=diamonds, aes(x=color)) + geom_bar()
```

For example: **jitter plots**
```{r}
ggplot(data=diamonds, aes(x=color, y = price/carat, colour=color)) + geom_jitter(alpha = 0.25)
```

### Layering your ggplot

Because of the step-wise nature of `ggplot` it is quite easy to overlay model fits, prediction lines, confidence intervals etc.
```{r}
baseplotcolor + geom_smooth()
```

If you wanted both the scatter plot and the smoothed fitted curve:
```{r}
baseplotcolor + geom_smooth() + geom_point()
```

Lets make the points smaller and more transparent so that fitted curves are more visible
```{r}
baseplotcolor + geom_point(size=1.5, alpha=0.25) + geom_smooth() 
```

### `ggplot` does maps very very well

Start by getting the `maps` library so that we can draw maps. Then create the map data for US states.
```{r}
library("maps") #Assuming this is installed, otherwise install first
USStates <- map_data("state")
```

We will use the `USArrests` data we have used before
```{r}
arrests <- USArrests #Create a copy
names(arrests) <- tolower(names(arrests)) #Converts column name to lower case
arrests$region <- tolower(rownames(USArrests)) #Same for row names

# Merge the map data with the arrests data based on region
arrestsbyregion <- merge(USStates, arrests, sort = FALSE, by = "region")
arrestsbyregion <- arrestsbyregion[order(arrestsbyregion$order), ]

# Plot a map, filling in the states based on murder rate
ggplot(data=arrestsbyregion, aes(x=long, y=lat, group=group, fill=murder)) + geom_polygon()
```

```{r}
ggplot(data=arrestsbyregion, aes(x=long, y=lat, group=group, fill=murder)) + geom_polygon() +
  scale_fill_gradient(low = "#FF0000", high = "#0000FF")
```

You can always save the last plot you generated using `ggsave`
```{r}
ggsave(filename = "mychart.jpg")
```

> **ggplot2** is very adaptable to variety of situations. I leave it to you to explore it further.

## `dplyr` Package

> `dplyr` package has very useful tools for dealing with data frames. It is one of most downloaded and used packages in R ecosystem today.

As usual, we start by loading the package (assuming it has been installed already). We will use the familiar `nycflights13` dataset to test out the package.
```{r}
library("dplyr"); library("nycflights13")
```
**Key Functions**: for each basic verb of data manipulation:

- filter() (and slice())
- arrange()
- select() (and rename())
- distinct()
- mutate() (and transmute())
- summarise()
- sample_n() and sample_frac()

**Filter for Selecting Rows**
```{r}
filter(flights, month == 1, day == 1)  #Criteria Combined using AND
filter(flights, month == 1 | month == 2) #Using OR
```

**Slice for Selecting Rows by Position**
```{r}
slice(flights, 1:10)
```

**Arrange Rows Using arrange()**
```{r}
arrange(flights, year, month, day) #Sorting in Ascending Order
arrange(flights, desc(arr_delay)) #Descending Order
```

**Select Columns With select()**
```{r}
select(flights, year, month, day)
select(flights, year:day) #Inclusive of limits
select(flights, -(year:day)) #Excluding all columns named
```

**Renaming Columns using select()**
```{r}
rename(flights, tail_num = tailnum)
```

**Extract Distinct/Unique Rows**
```{r}
distinct(select(flights, tailnum))
distinct(select(flights, origin, dest)) #All unqiue routes
```

**Adding New Columns With mutate()**
```{r}
mutate(flights, gain = arr_delay - dep_delay, speed = distance / air_time * 60)
mutate(flights,
  gain = arr_delay - dep_delay,
  gain_per_hour = gain / (air_time / 60) #You can refer to variables you just created
)
transmute(flights,
  gain = arr_delay - dep_delay,
  gain_per_hour = gain / (air_time / 60) #Only keep the new variables
)
```

**Summarise Values with summarise()**
```{r}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE))
```

**Randomly Sample Rows With sample_n() and sample_frac()**
```{r}
sample_n(flights, 10) #Fixed number of samples
sample_frac(flights, 0.01) #Fixed fraction of samples
```

### Grouped Operations in `dplyr` Using group_by()
```{r}
by_tailnum <- group_by(flights, tailnum)
delaydata <- summarise(by_tailnum,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE))
filtereddelay <- filter(delaydata, count > 20, dist < 2000)

ggplot(filtereddelay, aes(dist, delay)) +
  geom_point(aes(size = count), alpha = 1/2) + geom_smooth()
```

`dplyr` provides aggregating functions like: `n()` - number of observations, `n_distinct()` - nunber of distinct observations, `first()`, `last()`, `nth()` - self explanatory.

How many different planes go to each destination:
```{r}
destinations <- group_by(flights, dest)
summarise(destinations,
  planes = n_distinct(tailnum),
  flights = n())
```

### Chaining Functions Together
`dplyr` provides the `%>%` operator (also known as **pipe** operator) to chain many functions together.
```{r}
flights %>%
  group_by(year, month, day) %>%
  select(arr_delay, dep_delay) %>%
  summarise(
    arr = mean(arr_delay, na.rm = TRUE),
    dep = mean(dep_delay, na.rm = TRUE)
  ) %>%
  filter(arr > 30 | dep > 30)
```

