---
title: "Linear Regression Overview"
author: "Sanjeev Kumar"
date: "September 14, 2021"
output: 
  html_document: 
    highlight: tango
    theme: united
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

> You would want to keep a local copy of the `RMD` file and edit the file for the class exercise today. Submit an edited copy of the file as your class exercise submission.

<hr>

<p style="background-color:Orange;">Important Note: We will be doing class exercise today in a team of two students each. Figure out who you are working with. Move around if you need to.</p>
<hr>


# Work Flow

- Step 0: Preparation
- Step 1: Running Linear Regression and Evaluating the Results
- Step 2: Making Predictions
- Step 3: Evaluating the Model Performance
- Step 4: Iterate to Find Best Model


# Step 0: Preparation

Ensure that the data file is in the same directory as this .RMD file. Make that directory the working directory. `Session > Set Working Directory > To Source File Location`

## Read and Explore Data

```{r}
# Reading the data file 
Advertising <- read.csv("Advertising2.csv")
str(Advertising)

# Explore the data
if(!"GGally" %in% installed.packages()){install.packages("GGally")}
library(GGally)
ggpairs(Advertising)
```

## Data Cleaning

```{r}
# Start without Instagram (Factor) variable. 
# All x's are numeric variables 
Advertising
Ad_small <- Advertising[,-4] #Without instagram
Ad_small
```

##  Split Data into Training set and Validation set 

```{r}
#70-30 split 
set.seed(567) #All your partioins for this class use this seed. 
nrow(Ad_small)
train_set <- sample(1:nrow(Ad_small), 0.7*nrow(Ad_small))

#Training set 
tr <- Ad_small[train_set, ]
x_tr <- Ad_small[train_set,-4] #x only 
y_tr <- Ad_small[train_set,4] # y onle

#Validation set 
val <- Ad_small[-train_set, ]
x_val <- Ad_small[-train_set,-4]  #x only 
y_val <- Ad_small[-train_set,4]   #y only

```

# Step 1: Running Linear Regression

```{r}
# Simple linear regression: Sales = b0 + b1*TV 

m1 <- lm(Sales ~ TV, data =tr) #lm: linear model 

# Regression output 
summary(m1)

# Other regression outputs 
ls(m1) 
m1$coefficients
m1$fitted.values #est. y-values for the training set. 


# Residual plots 
plot(m1)

# Four plots in one single page 
par(mfrow=c(2,2))
plot(m1)

par(mfrow=c(1,1)) #resetting plot space back

# Plotting Regression Line
# x= TV, y = Sales
# line = regression line 
ggplot(tr) + 
  geom_point(aes(x=TV, y=Sales), color="blue") + 
  geom_smooth(aes(x=TV, y=Sales), method = "lm")+ 
  ggtitle("Scatter plot")


# x1= fitted.value (yhat) y= Sales (actual)
ggplot(data=tr) + 
  geom_point(aes(x= m1$fitted.values, y=Sales), color="blue") + 
  labs(title="Predicted vs. Actual", x ="Predicted y", y = "Actual y") 
```

# Step 2. Making predictions 

```{r}
# Apply m1 to validation data to generate prediction
# Point forecasts only 
predict(m1, x_val) #R automatically reads TV variable only. 


# Predicting the average value for given x 
# Confidence intervals: Estimates Average y values 
head(predict(m1, x_val, interval = "confidence", level = 0.95))


# Store point prediction to a vector 
pred1 <- predict(m1, x_val) 

```

#  Step 3. Evaluating the model performance   

```{r}
# Computing algorithm performance 
if(!"caret" %in% installed.packages()){install.packages("caret")}
library(caret)

# postResample (caret package): RMSE, R^2, MAE 
postResample(pred1, y_val)
m1_result <- postResample(pred1, y_val)


# Put them in the data frame for later comparisons 
result<- data.frame(m1 = round(m1_result,3))
result


#Want to compute MAPE (Mean Absolute Percentage Error) 
if(!"Metrics" %in% installed.packages()){install.packages("Metrics")}
library(Metrics)
mape(y_val, pred1) #mape(actual y, prediction)

```

# Step 4: Iterate to Find Best Model

## Full regression model 

> Do the class exercise below - fill the code for the three Steps. Out goal is to get the `result` object for the full model.

```{r}
#Exercise
# Run the full Regression Model  (m2) 
#  m2 <- lm(Sales ~ TV+ Radio + Newspaper, data = tr) 
# Examine Regression Outputs 
# Make predictions using validating data 

#  Step 1. Running linear regression and evaluating the results 

#  Step 2. Making predictions 

#  Step 3. Evaluating the model performance   
#          Put postResample result in the object, result     

```

## Smaller Regression Model

```{r}

# Model 3 (m3)  Sales ~ Tv+ Radio 

#  Step 1. Running linear regression and evaluating the results 

m3 <- lm(Sales ~ TV + Radio, data = tr)
summary(m3)


#  Step 2. Making predictions 

pred3 <- predict(m3, x_val)


#  Step 3. Evaluating the model performance   
#          Put postResample result in the object, result     

m3_result <- postResample(pred3, y_val)
result<- data.frame(result, m3 = round(m3_result,3))
result 
```

## Including dummy variable(s) 


```{r}
#Including dummy variables
#Creating Train and Validation Set for data with dummy variable

#70-30 split 
Ad_full <- read.csv("Advertising2.csv")
str(Ad_full)

#Train set 
tr <- Ad_full[train_set, ]
x_tr <- Ad_full[train_set,-5]
y_tr <- Ad_full[train_set,5]

#Validation set
val <- Ad_full[-train_set, ]
x_val <- Ad_full[-train_set,-5]  
y_val <- Ad_full[-train_set,5]  

head(tr)

#  Step 1. Running linear regression and evaluating the results 

md1 <- lm(Sales~., tr)
summary(md1)

#  Step 2. Making predictions 

pred4 <- predict(md1, x_val)

#Creating predict interval 
predict(md1, x_val, interval ="prediction", level= 0.95)


#  Step 3. Evaluating the model performance   
#          Put postResample result in the object, result     

# Evaluating model performance
postResample(pred4, y_val)
md1_result <- postResample(pred4, y_val)

result<- data.frame(result, md1 = round(md1_result,3))
result
```


## Interaction variables

```{r}
# Instagram*TV: Captures the interaction between Instagram and TV. 

# InstagramYes =1,    InstagramYes*TV = 1*TV = TV
# InstagramNo (InstagramYes=0)    InstagramYes*TV = 0*TV = 0

md2 <- lm(Sales~ TV + Newspaper + Radio + Instagram + Instagram*TV, tr)
summary(md2)
```

> Class Exercise - add model performace metrics for the `md2` model to the `result` object.

## Polynomial regression

```{r}
# I(polynomial math expression)
md3 <- lm(Sales ~ TV + I(TV ^ 2)  +Instagram, data = tr)
summary(md3)

```

> Class Exercise - add model performace metrics for the `md3` model to the `result` object.

# Conclusion: Best Model

> Class Exercise - which is the best Linear Regression model among all that we studied today?
