---
title: "w2_notes_regression"
author: "Ella.Li"
date: "3/15/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Linear Regression

## read, calean data
```{r}
adv <- read.csv("Advertising2.csv")

adv$Instagram <- as.factor(adv$Instagram)
str(adv)
summary(adv)
```
## split into test and train data
```{r}

set.seed(12345)
test_set <- sample(1:nrow(adv), 0.4*nrow(adv)) # take random sample, choose 40% data from all 1:200 rows
adv_test <- adv[test_set, ]
adv_train <- adv[-test_set, ] # data not in the test set

```

## build model

### simple model
```{r}

# response/y = sales, predictor/x = TV + Radio + Newspaper
m_1 <- lm(Sales ~ TV + Radio + Newspaper, data = adv_train)  # only use train data
summary(m_1)



```
(predicted) Sales = 2.988 + 0.0448*TV + 0.1967*Radio (no Newspaper, not significant)

### HOW TO HANDLE CATEGORICAL VARIABLES? -- Dummy variable model

```{r}

m2 <- lm(Sales ~ ., data= adv_train)
summary(m2)
```
### interaction variable model

```{r}

m3 <- lm(Sales ~ . + TV*Instagram, data = adv_train)
summary(m3)
```
WE can find that tv*ins campaign will decrease the slope of relationship of adv and sales.

### Step model (find best model)
```{r}
# starting from a big model: m3
mstep <- step(m3, direction = "backward")
summary(mstep)

```
the last result will give back only the significant predictors needed for prediction model.(here is TV and Radio)
