---
title: "TO_628_NOTE_1"
author: "Ella.Li"
date: "3/8/2022"
output:
  html_document: 
    toc: true
    toc_depth: 2
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Just starting

We are starting right now. **wish us luck!**

```{r}
# Just writing some code
print('Hello World!')

```
## Import data

we will start import data

```{r, cache=TRUE}
# import data
nyc <- read.csv("nycflights.csv")

# look structure
str(nyc)
# look summary of data
summary(nyc)



```

## clean data

Lets clean data now. *clean data is important*

```{r}
# delete the column called "x"
nyc$X <- NULL  # delete this column by assign NULL to it
nyc$year <- NULL

str(nyc)
```

### sepcial operaters

* NA means missing value;
* NULL means nothing (NULL represents the null object in R. NULL is used mainly to represent the lists with zero length, and is often * returned by expressions and functions whose value is undefined. as. null ignores its argument and returns the value NULL .);
* Inf means infinite;
* NaN means not a numer;

### convert CHR to a Factor (to go into the each uniq data)
```{r}
nyc$dest <- as.factor(nyc$dest)
nyc$origin <- as.factor(nyc$origin)
nyc$tailnum <- as.factor(nyc$tailnum)
nyc$carrier <- as.factor(nyc$carrier)

# look summary of data
summary(nyc)
```


# HW 1-PRACTICE
## DATA EXPLORATION

> How many airlines?

```{r}
number_of_airlines <- nlevels(nyc$carrier) # nlevels
number_of_airlines  # will only get result after factorize, always check the summary() to see structure

```
Number of airlines leaving NYC in 2013 was : **`r number_of_airlines`**.  #inline result

> Number of flights by the airline with code "OO"?

```{r}
nyc_oo <- nyc[nyc$carrier == "OO" , ]

```
Number of flights by aireline OO in 2012 was: `r nrow(nyc_oo)`

> How many flights where there by United Airlines (code: UA) on Jan 12th 2013?

```{r}
nyc_UA_JAN_12 <- nyc[(nyc$carrier == "UA") & (nyc$month == 1) & (nyc$day == 12) , ] # here no column just means every column
# here, int should == int type, but  == str will also work because R will auto transfer to same type
nrow(nyc_UA_JAN_12)
```
Number of flights by United Airlines (code: UA) on Jan 12th 2013 was: `r nrow(nyc_UA_JAN_12)`

> Shoetest flight time?

```{r}

min(nyc$air_time, na.rm = TRUE) # REMOVE NA
#rm(nycoo) # rm() function in R Language is used to delete objects from the memory


```

## Airline performance
Calculate average arrival delays by airline (Hint: look up the command tapply)

Draw a Bar Plot of Average Arrival Delays for all the Airlines (Hint: command for making a Bar Plot is simply barplot)

Which airline has the highest average arrival delay? Which airline has the smallest average arrival delay? Are there airlines that actually have negative average delay? Provide answer to this question in a text paragraph form using inline R code.

```{r}
airline_delay <- round(tapply(nyc$arr_delay, nyc$carrier, mean, na.rm=TRUE), 2) # do mean for each carrier's delay time using "tapply" and round to 0.00
airline_delay
barplot(airline_delay, las=2, cex.names = 0.7)  # las means rotation 90, cex.names changes the Size of name text
```

> Air Gain

calculate airgain values and keep it in a new column

```{r}
nyc$airgain <- nyc$dep_delay - nyc$arr_delay # creat a new column airgain
summary(nyc$airgain, na.rm=TRUE) # stast summary

airgain_each_carrier <- round(tapply(nyc$airgain, nyc$carrier, mean, na.rm=TRUE), 1) # for each carrier calculate mean of airgain and compare results
airgain_each_carrier

barplot(airgain_each_carrier, las=2, cex.names = 0.7) # plot

names(airgain_each_carrier[airgain_each_carrier == max(airgain_each_carrier)]) # filter the name of max mean.

```

